Machine Learning for Predictive Auto-Tuning

### 2011-12-11 paper/draft1.tex
paper

\section{Experiments}

@@

The core of any online auto-tuning algorithm (Fig.~\ref{fig:autotune_general}) is the search procedure used to explore the space of implementations.
Our implementation space includes 16000 (XXX) points, too many to search exhaustively.
\cite{cox+pinto:2011gcg} recommends a particular hyper-parameter implementation
based on empirical autotuning on several platforms, that achieves good
performance on a variety of GPUs from older-generation cards such as the
8600GT all the way to
current high-end cards such as the GTX 580 and C2070.
We call this configuration the {\em reference} implementation.

Additionally, \cite{cox+pinto:2011gcg} advocate a particular grid search over
the most promising part of the configuration space. This grid iterates over all
combinations of XXX, but leaves the value of XXX at. In our experiments, we call this algorithm the {\em grid} search procedure.

In this work, we consider a third search algorithm based on a local search heuristic.  This {\em hill-climing} algorithm starts from the reference implementation and randomly resamples $25\%$ of the implementation parameters randomly

XXX: Histogram (over problems) of gigaflops/sec of random, grid, random,
hill-climbing for each/selected platform.
This shows that our implementations and problems are stressing the cards,
that our implementation is pretty good, and that different problem sizes pose
different computational challenges.

XXX: Scatterplot of gigaflops vs. gigaflops / second (of best grid point)
should show that it's not just problem size that makes it possible to go fast.

XXX: histogram of speedups of grid, random, hill-climbing over reference, for
several hardware platforms. (GET ANOTHER -- macbook??)
Sometimes the reference implementation is optimal, in many cases it is far
from optimal: Which cases? What's wrong with the reference?

@@

\begin{figure*}
\centering
\includegraphics[scale=.42]{fig_ntrain_295.pdf}
\includegraphics[scale=.42]{fig_ntrain_480.pdf}
\caption{The effect of the invalid configuration score ($a$) and training set
size on simulated auto-tuning.
All candidates timed during the grid and
hill-climbing search procedures were used as training points, so the training
set sizes ranged from an average of 1500 (10 problem configurations) to 30
thousand (200 problem configurations).
Training on 10 or 25 configurations was useful (higher mean than the reference),
but not as useful as training on 50 or more configurations.
The results on the GTX 295 suggest that
moderate values of $a$ between $0.25$ and $0.75$ might be
best, but $a$ had no significant effect on the GTX 480.
}
\label{fig:fig_ntrain}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[scale=.42]{fig_gflop_scatter_295.pdf}
\includegraphics[scale=.42]{fig_gflop_scatter_480.pdf}
\caption{Computation speed for novel problem configurations using predictive
vs. empirical code specialization.
The scatterplots are roughly symmetric about the axis of equality (main diagonal)
with occasional outliers, indicating that for most problem configurations the
predictive approach gives as good a solution as the empirical approach.
The predictive approach typically requires about 0.1 seconds to suggest an
implementation, whereas the empirical approach requires about 2 minutes.
}
\label{fig:fig_gflop_scatter}
\end{figure*}

>>>>

\section{Experiment Setup}

@@

The core of any online auto-tuning algorithm (Fig.~\ref{fig:autotune_general}) is the search procedure used to explore the space of implementations.
Our implementation space includes 16000 (XXX) points, too many to search exhaustively.
\cite{cox+pinto:2011gcg} recommends a particular hyper-parameter implementation
based on empirical autotuning on several platforms, that achieves good
performance on a variety of GPUs from older-generation cards such as the
8600GT all the way to
current high-end cards such as the GTX 580 and C2070.
We call this configuration the {\em reference} implementation.
% XXX: what is reference?  Why was it chosen?

Additionally, \cite{cox+pinto:2011gcg} advocate a particular grid search over
the most promising part of the configuration space. This grid iterates over all
combinations of XXX, but leaves the value of XXX at. In our experiments, we call this algorithm the {\em grid} search procedure.
The grid included $72$ points in addition to the reference implementation, for a total of $73$ points.
XXX: isn't 73 prime?? explain this so-called ``grid''.

In this work, we also consider a third search algorithm based on a local search
heuristic.  This {\em hill-climing} (HC) algorithm starts from the reference
implementation and resamples each of the parameters
of the current best implementation randomly with 
probability $0.25$ (keeping the current best setting with probability $0.75$).
On each hill-climbing iteration, if the speed of the newly sampled point is greater than the previous point, then it becomes the current point.
We show results for search variants HC25, HC50, and HC75, which correspond do hill-climbin for 25, 50, and 75 iterations respectively.

\subsection{Platforms and Implementation}

We evaluated our approach on several GPU devices:
XXX.

XXX Table showing theoretical peak FLOPS, theoretical peak bandwidth, core count, shared memory size, register size, etc.

\section{Results}

Figure~\ref{fig:speedup} shows the effectiveness of empirical auto-tuning is in this setting.  Averaging across the range of problem configurations in our study,
implementations discovered by empirical auto-tuning are on average about 50\% (XXX check) faster than the reference implementation.
XXX: recalculate with simpler mean speedup formula.
On the GTX 295 speedups over the reference implementation ranged from XXX to XXX with mean XXX.
On the GTX 480 speedups over the reference implementation ranged from XXX to XXX with mean XXX.
XXX: Consider table for these numbers.
XXX: Put in some timings of mixed \& matched all-stars to show that indeed the reference was a good general baseline, and in fact the auto-tuned models are special-purpose.

Figure~\ref{fig:speedup} also shows that random hill-climbing is at least
competitive and often better than the hand-chosen grid approach of ~\citet{pinto+cox:2011gcg}.  Figure~\ref{fig:hcX} shows how the number of hill-climbing iterations
affects the quality of the auto-tuned model.
The quicker HC25 approach is clearly inferior but the quality of the 75 iteration search (HC75) is statistically similar to the shorter HC50, suggesting that XXX.
Still, HC75 is not always better than the grid, so evidently neither search algorithm is perfect.

The previous figures establish that auto-tuning on a problem-by-problem basis can achieve good performance improvements over a high-quality reference baseline, but they do not show how long it takes to find these implementations.  A search of 75 iterations took an average of about 120 seconds XXX on a fast computer, because it requires compiling 75 CUDA kernels, and evaluating up to 750 filterbank correlations to get reliable timing information for each candidate implementation.

Our contrib

@@

\begin{figure}
\centering
\includegraphics[scale=.42]{fig_ntrain_295.pdf}
\includegraphics[scale=.42]{fig_ntrain_480.pdf}
\caption{The effect of the invalid configuration score ($a$) and training set
size on simulated auto-tuning.
All candidates timed during the grid and
hill-climbing search procedures were used as training points, so the training
set sizes ranged from an average of 1500 (10 problem configurations) to 30
thousand (200 problem configurations).
Training on 10 or 25 configurations was useful (higher mean than the reference),
but not as useful as training on 50 or more configurations.
The results on the GTX 295 suggest that
moderate values of $a$ between $0.25$ and $0.75$ might be
best, but $a$ had no significant effect on the GTX 480.
}
\label{fig:fig_ntrain}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=.42]{fig_gflop_scatter_295.pdf}
\includegraphics[scale=.42]{fig_gflop_scatter_480.pdf}
\caption{Computation speed for novel problem configurations using predictive
vs. empirical code specialization.
The scatterplots are roughly symmetric about the axis of equality (main diagonal)
with occasional outliers, indicating that for most problem configurations the
predictive approach gives as good a solution as the empirical approach.
The predictive approach typically requires about 0.1 seconds to suggest an
implementation, whereas the empirical approach requires about 2 minutes.
}
\label{fig:fig_gflop_scatter}
\end{figure}
